Supplement to: Library Learning in Neurally-Guided Bayesian Program Induction

### 2018-10-26 nips_2018.tex
more writing

\title{Library Learning in Neurally-Guided Bayesian Program Learning}

@@

\begin{table}[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=3pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=4pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_0$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

@@

Our \system (ECC, for \textbf{E}xplore/\textbf{C}ompress/\textbf{C}ompile)
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Explore} step takes a given set of \textbf{tasks}, typically
several hundred, and explores the space of programs, searching for compact programs that solve these tasks,
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during exploration, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions:%% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
\noindent\\\textbf{Contributions.} (1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:
(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
  the space of  programs.

@@

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
highlighted in {\orange{orange}},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
\end{figure}
\begin{wrapfigure}{R}{0.6\textwidth}\vspace{-1cm}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.0cm}(\textbf{Explore})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.0cm}\textbf{(Explore)}}

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}\vspace{-1cm}
\end{wrapfigure}
To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{8pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{94\%} & 
  79\% &71\%&35\%&62\%&37\%\\
  Solve time&  88s&39s&11s&35s&44s&20s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&49s % FIXME: 65 in the table but 49 on the graph? 65s
  &38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.

%For each domain,
We are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
Tbl.~\ref{baselineComparisons}
compares our model against these alternatives.
We consistently
improve on the baselines,
and also find that lesioning the recognition model
impairs the convergence of the algorithm,
causing it to hit a lower `plateau' after which it stops solving new tasks, following an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
This lowered `plateau' supports a view of the recognition model as a way of amortizing the cost of search.

@@

@@

\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work. Funding from NSF GRFP and the MUSE program (Darpa grant FA8750-14-2-0242).

>>>>

\title{Library Learning in Neurally-Guided Bayesian Program Induction}

@@

\begin{table}[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=3pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=4pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_0$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$)}}
        \end{tabular}
        &
        \hspace{0.1cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

@@

Our \system (ECC, for \textbf{E}xplore/\textbf{C}ompress/\textbf{C}ompile)
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Explore} step takes a given set of \textbf{tasks}, typically
several hundred, and explores the space of programs, searching for compact programs that solve these tasks,
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during exploration, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference,stuhlmuller2013learning}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We find EC motivating, and go beyond it and other prior work through the following contributions:%% \begin{itemize}
%%   \setlength\itemsep{0}
%%   \item We show how to learn-to-learn programs in an expressive Lisp-like programming language;
%%   \item We give an algorithm for learning DSLs, built on a formalism known
%%     as Fragment Grammars~\cite{tim}; and 
%%   \item We give a hierarchical Bayesian framing of the problem that allows a
%%     joint inference of the DSL and recognition model.
%% \end{itemize}
%% \noindent\\\textbf{Contributions.} 
(1) We show how to learn-to-learn programs in an expressive Lisp-like programming language, including conditionals, variables, and higher-order recursive functions;
(2) We give an algorithm for learning DSLs,
built on a formalism known as Fragment Grammars~\cite{tim};
and (3) We give a hierarchical Bayesian framing of
the problem that
allows joint inference of the DSL and neural recognition model.

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta,polikarpova2016program}.
We have two reasons:
(1) A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
(2) Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
  the space of  programs.

@@

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure,
highlighted in {\orange{orange}},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}\label{fragmentExample}
\end{figure}
\begin{figure}%{R}{0.6\textwidth}\vspace{-1cm}
\centering  \begin{minipage}{0.9\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE For each task $x\in X$, set  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hfill(\textbf{Explore})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (see Sec.~\ref{recognitionSection})\hfill(\footnotesize{\textbf{Compile}})
        \STATE For each task $x\in X$, set $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hfill\textbf{(Explore)}}

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  (see Sec.~\ref{grammarInductionSection})\footnotesize{\hfill\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\end{figure}
To define the prior distribution over $(\mathcal{D},\theta})$, we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
and place a symmetric Dirichlet prior over the weight vector $\theta$.
%% , $  \probability[\theta|\mathcal{D}] = \text{Dir}(\theta|\alpha)$, where
 %% $\alpha$ is a concentration parameter controlling the smoothness of the prior over $\theta$.
%Alg.~\ref{grammarInductionAlgorithm} specifies the DSL induction algorithm.

%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
For added robustness,
we interleave an extra program search step (Exploration) before training the recognition model,
and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Exploration step.

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.

\begin{wraptable}{r}{0.55\textwidth}
\vspace{-10pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{94\%} & 
  79\% &71\%&35\%&62\%&37\%\\
  Solve time&  88s&39s&11s&35s&44s&20s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&49s % FIXME: 65 in the table but 49 on the graph? 65s
  &38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-1cm}\label{baselineComparisons} \end{wraptable}
%For each domain,
We are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
Tbl.~\ref{baselineComparisons}
compares our model against these alternatives.
We consistently
improve on the baselines,
and also find that lesioning the recognition model
impairs the convergence of the algorithm,
causing it to hit a lower `plateau' after which it stops solving new tasks, following an initial spurt of learning (Fig.~\ref{learningCurves}) -- without the neural network, search becomes intractable.
This lowered `plateau' supports a view of the recognition model as a way of amortizing the cost of search.

@@

 \section{Limitations of EC$^2$}

 We have presented an algorithm that, starting from relatively little,
 learned to solve nontrivial program synthesis problems in three
 different domains.  However, our system was also given (and also,
 critically \emph{needs}) a corpus of training tasks to solve for each
 domain.  Is constructing (or curating) corpra of tasks any easier or
 better than hand-engineering DSLs?  In the immediate future, we
 expect hand-engineering of DSLs to largely continue, especially in
 domains where humans have strong intuitions about the underlying
 system of domain specific concepts, like text editing. However, if
 program induction is to become a standard part of the AI toolkit,
 then, in the long-term, we need to build agents that autonomously
 acquire the domain specific knowledge needed to navigate a new
 domain.  So, through the lens of program synthesis, EC$^2$ carries
 the limitation that it requires a high-quality corpus of training
 tasks; but, for the program-induction approach to AI, this limitation
 is natural.

 The particular implementation of EC$^2$ developed here
 inherits a key limitation from a specific arbitrary engineering decision:
 our enumerative program search strategy cannot possibly
 scale to large bodies of code, even with a good recognition model --- but one could in principle replace
 our enumerative approach with stochastic search~\cite{schkufza2013stochastic,DBLP:books/daglib/0070933},
 constraint solving~\cite{solar2008program}, type-directed synthesis~\cite{polikarpova2016program},
 or other techniques from the program synthesis community.

@@

\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work, and for funding from the NSF GRFP, the MUSE program (Darpa grant FA8750-14-2-0242),
and an AWS ML Research Award.

### 2018-10-26 supplement.tex
more writing

\title{Supplement to: Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning}

@@

\author{
  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay
  \AND
  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

>>>>

\title{Supplement to: Library Learning in Neurally-Guided Bayesian Program Induction}

@@

\author{
  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\\\texttt{lucasem@mit.edu}\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay\\\texttt{mathsm@mit.edu}
  \AND
  Armando Solar-Lezama\\MIT\\\texttt{asolar@csail.mit.edu}\And Joshua B. Tenenbaum\\MIT\\\texttt{jbt@mit.edu}
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}
